---
title: "DSA3361_RMarkdown_2-2"
output:
  pdf_document: default
  html_document: default
date: "`r Sys.Date()`"
---
```{r, hide = TRUE}
#install.packages("tinytex")
library(tinytex)
```

1. Find the mean and standard deviation of the population, pop, listed as  
pop <- c(3,5,6,6,8,11,15,15,19,20)  
Visualize the population using a stripchart().

```{r}
pop <- c(3,5,6,6,8,11,15,15,19,20)  
stripchart(pop,
           pch = 19,
           method = "stack",
           at = 0,
           main = "Population")
```
2. Simulate sampling by taking  
- 5 samples of size 3 without replacement.   
- 5 samples of size 3 with replacement.   
- 10000 samples of size 10 without replacement.   
- 10000 samples of size 10 with replacement.   

```{r}
set.seed(4224)
N <- 5
for (i in 1:N) {
  s1 <- sample(pop, size = 3, replace = TRUE)
  print(s1)
}
```
3. Simulate the sampling distribution X-bar, by taking 10000 samples of size 4
- Use a histogram to plot your results.  
- Compute the mean and standard error of the sampling distribution.   
- Compare those values to the population mean and standard deviation.  

```{r}
set.seed(4224)
N <- 1e4
samples_mean_rep <- numeric(N)
for (i in 1:N) {
  samples_mean_rep[i] <- mean(sample(pop, size = 4, replace = TRUE))
}
hist(samples_mean_rep, main = "Sample mean distribution with rep", col = "green")

samples_mean_norep <- numeric(N)
for (i in 1:N) {
  samples_mean_norep[i] <- mean(sample(pop, size = 4, replace = FALSE))
}
hist(samples_mean_norep, main = "Sample mean distribution w/o rep", col = "blue")


```

```{r}
mean_x_rep <- mean(samples_mean_rep)
mean_x_norep <- mean(samples_mean_norep)
mean_pop <- mean(pop)
print(c(mean_x_rep, mean_x_norep, mean_pop))

print(c(sd(samples_mean_rep), sd(samples_mean_norep), sd(pop) / sqrt(4)))

```
